---
title: "NCVROC v0.1 — Introduction"
author: "NCVROC"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{NCVROC-intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

NCVROC develops short item-based screening scales through combinatorial
item-set selection, ROC-based evaluation, and nested cross-validation.
Given a set of questionnaire items and a binary outcome, NCVROC identifies
which small subset of items best predicts the outcome using a simple sum
score.

Key assumptions:

- **Higher sum score = higher probability of a positive outcome.** Users
  must reverse-code items beforehand if needed.
- The cutoff rule is `score >= cutoff` (fixed direction).
- v0.1 does not handle missing values; any `NA` in the data causes an error.

## Simulated data

We create a dataset with 200 rows and 6 items scored 0–2. Items q1–q3
carry signal for the outcome; q4–q6 are noise.

```{r}
library(NCVROC)

set.seed(42)
n <- 200

# Signal items (q1–q3): higher values increase probability of positive outcome
q1 <- sample(0:2, n, replace = TRUE)
q2 <- sample(0:2, n, replace = TRUE)
q3 <- sample(0:2, n, replace = TRUE)

# Noise items (q4–q6): independent of outcome
q4 <- sample(0:2, n, replace = TRUE)
q5 <- sample(0:2, n, replace = TRUE)
q6 <- sample(0:2, n, replace = TRUE)

# Outcome: probability depends on sum of signal items
signal_sum <- q1 + q2 + q3
prob <- plogis(-2 + 0.5 * signal_sum)  # logistic transform
y <- rbinom(n, 1, prob)

d <- data.frame(y = y, q1 = q1, q2 = q2, q3 = q3, q4 = q4, q5 = q5, q6 = q6)

table(y = d$y)
```

About `r sum(d$y)` positive cases out of `r nrow(d)`.

## Exhaustive ROC evaluation

First, we enumerate all 1- and 2-item combinations and evaluate their ROC
performance on the full data. This is useful for quick screening but the
metrics are **apparent (in-sample) estimates**.

```{r}
exhaustive <- exhaustive_sum_roc(d, "y", paste0("q", 1:6),
                                  max_items = 2, progress = FALSE)
head(exhaustive, 10)
```

The top models by AUC are dominated by signal items (q1–q3).

## Nested cross-validation

For validated performance estimates, we use nested cross-validation.
The outer loop estimates generalization performance; the inner loop
selects the best model without leaking test information.

```{r}
result <- nested_sum_roc(
  data       = d,
  outcome    = "y",
  items      = paste0("q", 1:6),
  max_items  = 2,
  outer_k    = 3,
  inner_k    = 2,
  seed       = 123,
  progress   = FALSE
)
```

### Summary

```{r}
summary(result)
```

The mean AUC ± SD gives a cross-validated estimate of predictive
performance.

### Selected model frequency

```{r}
result$selected_model_frequency
```

This shows how often each item set was selected as the best model across
outer folds.

### Plot: model selection frequency

```{r, fig.width = 7, fig.height = 4}
plot(result, which = "selection")
```

### Plot: per-fold AUC

```{r, fig.width = 6, fig.height = 4}
plot(result, which = "auc")
```

The horizontal reference line at AUC = 0.5 marks chance-level performance.

## Final scale on full data

After cross-validation confirms that the selected items have predictive
value, `fit_final_sum_scale()` fits the top models on the complete dataset.
This gives the final candidate scales to recommend for future use.

**Important:** The metrics returned by `fit_final_sum_scale()` are
**apparent in-sample estimates**, not cross-validated. They will be
optimistically biased. Use `nested_sum_roc()` for validated performance
estimates.

```{r}
final <- fit_final_sum_scale(d, "y", paste0("q", 1:6),
                              max_items = 2, progress = FALSE)
head(final, 5)

# Verify the performance_type attribute
attr(final, "performance_type")
```

## Summary of apparent vs. nested CV performance

| Function              | Performance estimate        | Use case                        |
|-----------------------|-----------------------------|---------------------------------|
| `exhaustive_sum_roc()` | Apparent (in-sample)       | Quick screening, exploration    |
| `nested_sum_roc()`    | Nested cross-validated      | Validated performance estimation |
| `fit_final_sum_scale()` | Apparent (in-sample)      | Final scale on full data        |

Always use `nested_sum_roc()` when reporting performance. Use
`fit_final_sum_scale()` only after validation, to create the final
recommended item sets.
